---
title: "data.table Exercises - Solution"
author: "Janek Thomas"
date: "25. April 2016"
output: html_document
---

1. Load the `mpg` dataset from the package `ggplot2`. Check which class the dataset has and 
    convert it to a `data.table`.
    
```{r}
library(data.table)
data(mpg, package = "ggplot2")
class(mpg)
mpg = data.table(mpg)
class(mpg)
```

2. Set manufacturer and class as keys for the dataset. Then, select the subset of every nissan and
audi compact.

```{r}
setkey(mpg, manufacturer, class)
# alternative:
# setkeyv(mpg, c("manufacturer", "class"))
mpg[.(c("nissan", "audi"), "compact")]

```

    
3. Calculate the frequencies in which each combination of manufacturer and class is 
    observered.

```{r}
mpg[, .N, by = .(manufacturer, class)]
```

4.  Which manufacturer builds the cars with the most engine displacement on average?

```{r}
mpg[, .(meanDispl = mean(displ)), by = .(manufacturer)][order(meanDispl, decreasing = TRUE)]
```

    
5. Calculate a new variable that is the deviation of the displacement from the mean 
    displacement for its class.
    
```{r}
mpg[, displDev := displ - mean(displ), by = .(class)]
mpg
```

    
6.  Calculate the average city miles per gallon relativ to the number of cylinders for every dodge, 
    grouped by the year the car was build.
    
```{r}
mpg["dodge", .(relMpg = mean(cty / cyl)), by = .(year)]
```

    
7.  Which car from year 1999 has the biggest difference between city and highway miles per gallon?

```{r}
mpg[, diff := abs(cty - hwy)]
mpg[year == 1999][order(diff, decreasing = TRUE)]
```

    
8.  Predict the displacement with a linear model from the year, number of cylinders, city and 
    highway miles per gallon. Use a seperate model for every combination of manufacturer and class
    and save your predictions as a new variable in the dataset.

   
```{r}
mpg[, predDisp := predict(lm(displ ~ year + cyl + cty + hwy)), by = .(manufacturer, class)]
mpg
```

